---
title: "Fantasy vs. Reality"
author: "Jake Berberian"
output: pdf_document
header-includes:
  - \usepackage{xcolor}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

```{r library, include = FALSE}
library(tidyverse)
library(kableExtra)
```

# Data Import and Clean

```{r data, include = FALSE}
pbp <- vroom::vroom("~/Downloads/pbp.csv")
```


```{r game_scores}
scores <- pbp %>% 
  group_by(game_id) %>% 
  summarize(n = n(),
            home_team = home_team[n],
            home_score = total_home_score[n],
            away_team = away_team[n],
            away_score = total_away_score[n]) %>% 
  select(-n) 

scores %>% 
  head(5) %>% 
  kbl(booktabs = TRUE) %>% 
  kable_styling(latex_options = "striped")
```

Columns that are important: 

- QB/RB/WR/TE
  - play_type
  - yards_gained
  - td_team
  - pass_touchdown
  - rush_touchdown
  - return_touchdown
- Kickers
  - field_goal_result
  - kick_distance
- D/ST
  - interception
  - fumble_forced
  - safety
  - fumble_lost
  - sack
  
## Skill Positions

Here's a tricky thing: *return_touchdown* counts all types of returns (interceptions, fumbles, punts, etc.). So we'll need to deal with that on our own because a fumble return TD is only worth 6 points to the D/ST, but a kickoff return TD is worth 12 (6 to the individual player, 6 to the D/ST).

PPR or non-PPR

```{r touchdowns}
tds_tos <- pbp %>% 
  group_by(game_id, posteam) %>% 
  summarize(pass_td = sum(pass_touchdown),
            rush_td = sum(rush_touchdown),
            fumble = sum(fumble_lost),
            int = sum(interception)) %>% 
  drop_na() %>% 
  mutate(ft_pts = 6*(pass_td + rush_td) - 2*fumble - 1*int) %>% 
  rename(score_team = posteam) %>% 
  select(game_id, score_team, ft_pts)

yardage <- pbp %>% 
  group_by(game_id, posteam, play_type) %>% 
  summarize(yrds = sum(yards_gained)) %>% 
  drop_na() %>% 
  filter(play_type %in% c("pass", "run", "qb_kneel")) %>% 
  mutate(ft_pts = case_when(
    play_type == "pass" ~ yrds * 0.14,
    play_type == "run" ~ yrds * 0.1,
    play_type == "qb_kneel"~ yrds * 0.1)) %>% 
  rename(score_team = posteam) %>% 
  ungroup() %>% 
  select(game_id, yrds, ft_pts, score_team) %>% 
  group_by(game_id, score_team) %>% 
  summarize(ft_pts = sum(ft_pts))


returns <- pbp %>% 
  filter(return_touchdown == 1) %>% 
  group_by(game_id, play_type, posteam, defteam) %>% 
  summarize(kick_ret = sum(play_type == "kickoff" | play_type == "punt"),
            def_ret = sum(play_type == "pass" | play_type == "run")) %>% 
  ungroup() %>% 
  mutate(score_team = case_when(
    kick_ret != 0 ~ posteam,
    def_ret != 0 ~ defteam)) %>% 
  ungroup() %>% 
  mutate(ft_pts = 6 * def_ret + 12 * kick_ret) %>% 
  group_by(game_id, score_team) %>% 
  summarize(ft_pts = sum(ft_pts))
```


All looks good: checked a few randomly selected games using ProFootballReference boxscores. 

## Kickers

```{r kickers}
fgs <- pbp %>% 
  filter(play_type == "field_goal" | play_type == "extra_point") %>% 
  select(game_id, posteam, play_type, field_goal_result, kick_distance) %>% 
  group_by(game_id, posteam) %>%
  summarize(fg19 = sum(field_goal_result == "made" & kick_distance < 20),
            fg29 = sum(field_goal_result == "made" & kick_distance > 20 & kick_distance < 30),
            fg39 = sum(field_goal_result == "made" & kick_distance > 30 & kick_distance < 40),
            fg49 = sum(field_goal_result == "made" & kick_distance > 40 & kick_distance < 50),
            fg50 = sum(field_goal_result == "made" & kick_distance > 49)) %>% 
  mutate(ft_pts = 3 * fg19 + 3 * fg29 + 3 * fg39 + 4 * fg49 + 5 * fg50) %>% 
  ungroup() %>% 
  select(game_id, posteam, ft_pts) %>% 
  rename(score_team = posteam)


pats <- pbp %>% 
  filter(play_type == "extra_point") %>% 
  select(game_id, posteam, extra_point_result) %>% 
  group_by(game_id, posteam) %>% 
  summarize(pats = sum(extra_point_result == "good")) %>% 
  mutate(ft_pts = 1 * pats) %>% 
  ungroup() %>% 
  select(-pats) %>% 
  rename(score_team = posteam)

kickers <- full_join(fgs, pats, by = c("game_id", "score_team")) %>% 
  mutate(ft_pts.x = replace_na(ft_pts.x, 0),
         ft_pts.y = replace_na(ft_pts.y, 0),
         ft_pts = ft_pts.x + ft_pts.y) %>% 
  select(-c(ft_pts.x, ft_pts.y))
```


## Defense and Special Teams

I don't think blocked punts count for additional points, so we won't include them.
Need to add two point conversions.

```{r dst}
turnovers <- pbp %>% 
  group_by(game_id, defteam) %>% 
  summarize(int = sum(interception),
            sack = sum(sack),
            sfty = sum(safety),
            fmbl = sum(fumble_lost)) %>% 
  drop_na() %>% 
  mutate(ft_pts = sack + 2*(int+sfty+fmbl)) %>% 
  rename(score_team = defteam) %>% 
  select(game_id, score_team, ft_pts)

points_allowed <- scores %>% 
  pivot_longer(cols = c(away_score, home_score), names_to = "team", values_to = "score") %>% 
  mutate(def_team = case_when(
    team == "home_score" ~ away_team,
    team == "away_score" ~ home_team)) %>% 
  rename(pts_allowed = score) %>% 
  select(game_id, def_team, pts_allowed) %>% 
  mutate(ft_pts = case_when(
    pts_allowed == 0 ~ 10,
    pts_allowed < 7 & pts_allowed > 0 ~ 7,
    pts_allowed < 14 & pts_allowed > 6 ~ 4,
    pts_allowed < 21 & pts_allowed > 13 ~ 1,
    pts_allowed < 28 & pts_allowed > 20 ~ 0,
    pts_allowed < 35 & pts_allowed > 27 ~ -1,
    pts_allowed > 34 ~ -4,
  )) %>% 
  rename(score_team = def_team) %>% 
  select(-pts_allowed)


blks <- pbp %>% 
  filter(field_goal_result == "blocked") %>% 
  group_by(game_id, defteam, field_goal_result) %>% 
  summarize(blk = sum(field_goal_result == "blocked")) %>% 
  mutate(ft_pts = 2 * blk) %>% 
  ungroup() %>% 
  rename(score_team = defteam) %>% 
  select(game_id, score_team, ft_pts)
```



## Fantasy Points: Total
 
```{r join}
fantasy <- yardage %>% 
  bind_rows(tds_tos) %>% 
  bind_rows(returns) %>% 
  bind_rows(kickers) %>% 
  bind_rows(turnovers) %>% 
  bind_rows(points_allowed) %>% 
  bind_rows(blks) %>% 
  group_by(game_id, score_team) %>% 
  summarize(ft_pts = sum(ft_pts)) 

fantasy %>% 
  head(10) %>% 
  kbl(booktabs = TRUE) %>% 
  kable_styling(latex_options = "striped")
```



We need to get ready for a join with the scores dataframe.

```{r prep_scores}
new_scores <- scores %>% 
  pivot_longer(cols = c(away_score, home_score), names_to = "team", values_to = "score") %>% 
  mutate(score_team = case_when(
    team == "home_score" ~ home_team,
    team == "away_score" ~ away_team)) %>% 
  select(game_id, score_team, score)

new_scores %>% 
  head(10) %>% 
  kbl(booktabs = TRUE) %>% 
  kable_styling(latex_options = "striped")
```


```{r join_scores}
final <- right_join(fantasy, new_scores, by = c("game_id", "score_team"))

final %>% 
  head(10) %>% 
  kbl(booktabs = TRUE) %>% 
  kable_styling(latex_options = "striped")
```


# Results time!

```{r results}
final_reveal <- final %>% 
  group_by(game_id) %>% 
  summarize(W_game = score_team[which.max(score)],
            W_ft = score_team[which.max(ft_pts)]) %>% 
  mutate(diff = W_game == W_ft) %>% 
  count(diff)

final_reveal %>% 
  kbl(booktabs = TRUE) %>% 
  kable_styling(latex_options = "striped")

final_reveal %>% 
  summarize(diff, prop = round(n / sum(n), 4)) %>% 
  kbl(booktabs = TRUE) %>% 
  kable_styling(latex_options = "striped")
```


# Some Brief Analysis

Let's look at the largest difference in fantasy points. 

```{r analysis}
analysis <- final %>% 
  group_by(game_id) %>% 
  summarize(W_score = max(score),
            L_score = min(score),
            W_ft = max(ft_pts),
            L_ft = min(ft_pts)) %>% 
  mutate(diff_score = W_score - L_score,
         diff_ft = W_ft - L_ft) 

analysis %>% 
  arrange(desc(diff_score)) %>% 
  head(10) %>% 
  kbl(booktab = TRUE) %>% 
  kable_styling(latex_options = "striped")

analysis %>% 
  arrange(desc(diff_ft)) %>% 
  head(10) %>% 
  kbl(booktab = TRUE) %>% 
  kable_styling(latex_options = "striped")
```





